---
title: "Part 2"
author: "Pedro Lins"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages
library(gapminder)
library(tidyverse)
library(ggplot2)
```

# Load your data

```{r}
# Upload data as a .csv file
ds = "openpowerlifting.csv" 
df = read_csv(ds)
head(df)
```

# Exploratory Data Analysis
# Lifters Analysis
The goal here is to create an app that people can query a lifter from the database and see his performance statistics. 

initially, lets evaluate the general formula for attempts index
if a lifter misses a lift, it shows as a negative value
if a lifter performed it, positive
so we want to count how many positives a lifter had and divide it by all attempts he did
this gives the probability of a lifter performing a good lift
this can be drilled down to each lift (squat, bench press, deadlift)
and by time (is it changing through time?)


```{r}
##################
# Attempts Index #
##################

#########
# Squat #
#########

Name1 = "Pedro Lins"
Comptype = "Raw"
# First squat
sq1 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Squat1Kg > 0)
colnames(sq1) = c("Squat1", "sq1")

  if(sq1[1,1] == TRUE){
    sq1idx = data.frame(c(1))
  } else {sq1idx = sq1[2,2] / (sq1[1,2] + sq1[2,2])
    }
colnames(sq1idx) = c("sq1idx")


# Second Squat
sq2 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Squat2Kg > 0)
colnames(sq2) = c("Squat2", "sq2")

if(sq2[1,1] == TRUE){
    sq2idx = data.frame(c(1))
  } else {sq2idx = sq2[2,2] / (sq2[1,2] + sq2[2,2])
    }
colnames(sq2idx) = c("sq2idx")


# Thrid Squat
sq3 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Squat3Kg > 0)
colnames(sq3) = c("Squat3", "sq3")

if(sq3[1,1] == TRUE){
    sq3idx = data.frame(c(1))
  } else {sq3idx = sq3[2,2] / (sq3[1,2] + sq3[2,2])
    }
colnames(sq3idx) = c("sq3idx")

# Fourth Squat will not get into the analysis since it is technically an outlier

sq = (sq1idx + sq2idx + sq3idx)/3
colnames(sq) = c("sq")
sq
```

```{r}
###############
# Bench Press #
###############

# First Bench Press
bp1 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Bench1Kg > 0)
colnames(bp1) = c("Bench1", "bp1")

  if(bp1[1,1] == TRUE){
    bp1idx = data.frame(c(1))
  } else {bp1idx = bp1[2,2] / (bp1[1,2] + bp1[2,2])
    }
colnames(bp1idx) = c("bp1idx")
bp1idx

# Second Bench Press
bp2 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Bench2Kg > 0)
colnames(bp2) = c("Bench2", "bp2")

  if(bp2[1,1] == TRUE){
    bp2idx = data.frame(c(1))
  } else {bp2idx = bp2[2,2] / (bp2[1,2] + bp2[2,2])
    }
colnames(bp2idx) = c("bp2idx")
bp2idx

# Thrid Bench Press
bp3 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Bench3Kg > 0)
colnames(bp3) = c("Bench3", "bp3")

  if(bp3[1,1] == TRUE){
    bp3idx = data.frame(c(1))
  } else {bp3idx = bp3[2,2] / (bp3[1,2] + bp3[2,2])
    }

colnames(bp3idx) = c("bp3idx")


# Fourth Bench press will not get into the analysis since it is technically an outlier

bp = (bp1idx + bp2idx + bp3idx)/3
colnames(bp) = c("bp")
bp
```

```{r}
############
# Deadlift #
############

# First Deadlift
dl1 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Deadlift1Kg > 0)
colnames(dl1) = c("Deadlift1", "dl1")

  if(dl1[1,1] == TRUE){
    dl1idx = data.frame(c(1))
  } else {dl1idx = dl1[2,2] / (dl1[1,2] + dl1[2,2])
    }
colnames(dl1idx) = c("dl1idx")
dl1idx

# Second Deadlift
dl2 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Deadlift2Kg > 0)
colnames(dl2) = c("Deadlift2", "dl2")

  if(dl2[1,1] == TRUE){
    dl2idx = data.frame(c(1))
  } else {dl2idx = dl2[2,2] / (dl2[1,2] + dl2[2,2])
    }
colnames(dl2idx) = c("dl2idx")
dl2idx

# Thrid Deadlift
dl3 = df %>%
  filter(Name == Name1) %>%
  filter(Equipment == Comptype) %>%
  count(Deadlift3Kg > 0)
colnames(dl3) = c("Deadlift3", "dl3")

  if(dl3[1,1] == TRUE){
    dl3idx = data.frame(c(1))
  } else {dl3idx = dl3[2,2] / (dl3[1,2] + dl3[2,2])
    }
colnames(dl3idx) = c("dl3idx")
dl3idx

# Fourth Deadlift will not get into the analysis since it is technically an outlier

dl = (dl1idx + dl2idx + dl3idx)/3
colnames(dl) = c("dl")
dl
```

```{r}
#################
# Overall Index #
#################

index = (sq + bp + dl)/3
sq_index = data.frame(sq1idx, sq2idx, sq3idx, sq)
sq_trans <- as.data.frame(t(sq_index))
sq_trans
bp_index = data.frame(bp1idx, bp2idx, bp3idx, bp)
bp_trans<- as.data.frame(t(bp_index))
bp_trans
dl_index = data.frame(dl1idx, dl2idx, dl3idx, dl)
dl_trans<- as.data.frame(t(dl_index))
dl_trans


Attempts = c(1,2,3,"Average")
lifteridx = c(sq, bp, dl, index)

df_lifter = data.frame(Attempts, sq_trans, bp_trans, dl_trans, lifteridx)
colnames(df_lifter) = c("Attempts", "Squat", "Bench", "Deadlift", "Index")
print(df_lifter)

p1 = ggplot(df_lifter, aes(y = Squat, x = Attempts))+
         geom_col()
p2 = ggplot(df_lifter, aes(y = Bench, x = Attempts))+
         geom_col()
p3 = ggplot(df_lifter, aes(y = Deadlift, x = Attempts))+
         geom_col()
p4 = ggplot(df_lifter, aes(y = Index, x = Attempts))+
         geom_col()

```

# Jump analysis
During the competition, lifters can only increase the weight for the next attempt. Because of that, lifters manage risk by starting with a weight that is more probable to lift and increase for the next according to previous result. Here, we analyse how the increase can impact competition performance. 

How much is too much of a jump so that the lifter might miss the attempt? 
How much is a safe weight increase for next attempt?

# Squat

```{r}
###############
# Squat Jumps #
###############

##############################
# Good attempts % comparison #
##############################

# When your second lift is GOOD, how much percentage is it from the first?
sq2good = df %>%
  filter(Name == Name1) %>%
  filter(Squat2Kg > 0)

sq2j1good = 100 * ((abs(sq2good[,12]) / abs(sq2good[,11]))-1)
mean(sq2j1good$Squat2Kg)

# When your third lift is GOOD, how much percentage is it from the second?
sq3good = df %>%
  filter(Name == Name1) %>%
  filter(Squat3Kg > 0)

sq3j2good = 100 * ((abs(sq3good[,13]) / abs(sq3good[,12]))-1)
mean(sq3j2good$Squat3Kg)
```

```{r}
#############################
# Bad attempts % comparison #
#############################

# When your second lift is BAD, how much percentage is it from the first?
sq2bad = df %>%
  filter(Name == Name1) %>%
  filter(Squat2Kg < 0)
  
sq2j2bad = 100 * ((abs(sq2bad[,12]) / abs(sq2bad[,11]))-1)
mean(sq2j2bad$Squat2Kg)

# When your third lift is BAD, how much percentage is it from the second?
sq3bad = df %>%
  filter(Name == Name1) %>%
  filter(Squat3Kg < 0)

sq3j2bad = 100 * ((abs(sq3bad[,13]) / abs(sq3bad[,12]))-1)
mean(sq3j2bad$Squat3Kg)
```

```{r}
###########################
# Attempts % when 3 for 3 #
###########################

sqgood = df %>%
  filter(Name == Name1) %>%
  filter(Squat1Kg > 0) %>%
  filter(Squat2Kg > 0) %>%
  filter(Squat3Kg > 0) 

sqjump1good = 100 * ((abs(sqgood[,12]) / abs(sqgood[,11]))-1)
sqjump2good = 100 * ((abs(sqgood[,13]) / abs(sqgood[,12]))-1)

mean(sqjump1good$Squat2Kg)
mean(sqjump2good$Squat3Kg)
mean(sqjump1good$Squat2Kg) + mean(sqjump2good$Squat3Kg)
```

# Bench Press
```{r}
#####################
# Bench Press Jumps #
#####################

##############################
# Good attempts % comparison #
##############################

# When your second lift is GOOD, how much percentage is it from the first?
bp2good = df %>%
  filter(Name == Name1) %>%
  filter(Bench2Kg > 0)

bp2j1good = 100 * ((abs(bp2good[,17]) / abs(bp2good[,16]))-1)
mean(bp2j1good$Bench2Kg)

# When your third lift is GOOD, how much percentage is it from the second?
bp3good = df %>%
  filter(Name == Name1) %>%
  filter(Bench3Kg > 0)

bp3j2good = 100 * ((abs(bp3good[,18]) / abs(bp3good[,17]))-1)
mean(bp3j2good$Bench3Kg)
```

```{r}
#############################
# Bad attempts % comparison #
#############################

# When your second lift is BAD, how much percentage is it from the first?
bp2bad = df %>%
  filter(Name == Name1) %>%
  filter(Bench2Kg < 0)
  
bp2j2bad = 100 * ((abs(bp2bad[,17]) / abs(bp2bad[,16]))-1)
mean(bp2j2bad$Bench2Kg)

# When your third lift is BAD, how much percentage is it from the second?
bp3bad = df %>%
  filter(Name == Name1) %>%
  filter(Bench3Kg < 0)

bp3j2bad = 100 * ((abs(bp3bad[,18]) / abs(bp3bad[,17]))-1)
mean(bp3j2bad$Bench3Kg)
```

```{r}
###########################
# Attempts % when 3 for 3 #
###########################

bpgood = df %>%
  filter(Name == Name1) %>%
  filter(Bench1Kg > 0) %>%
  filter(Bench2Kg > 0) %>%
  filter(Bench3Kg > 0) 

bpjump1good = 100 * ((abs(bpgood[,17]) / abs(bpgood[,16]))-1)
bpjump2good = 100 * ((abs(bpgood[,18]) / abs(bpgood[,17]))-1)

mean(bpjump1good$Bench2Kg)
mean(bpjump2good$Bench3Kg)
mean(bpjump1good$Bench2Kg) + mean(bpjump2good$Bench3Kg)
```

# Deadlift
```{r}
##################
# Deadlift Jumps #
##################

##############################
# Good attempts % comparison #
##############################

# When your second lift is GOOD, how much percentage is it from the first?
dl2good = df %>%
  filter(Name == Name1) %>%
  filter(Deadlift2Kg > 0)

dl2j1good = 100 * ((abs(dl2good[,22]) / abs(dl2good[,21]))-1)
mean(dl2j1good$Deadlift2Kg)

# When your third lift is GOOD, how much percentage is it from the second?
dl3good = df %>%
  filter(Name == Name1) %>%
  filter(Deadlift3Kg > 0)

dl3j2good = 100 * ((abs(dl3good[,23]) / abs(dl3good[,22]))-1)
mean(dl3j2good$Deadlift3Kg)
```

```{r}
#############################
# Bad attempts % comparison #
#############################

# When your second lift is BAD, how much percentage is it from the first?
dl2bad = df %>%
  filter(Name == Name1) %>%
  filter(Deadlift2Kg < 0)
  
dl2j2bad = 100 * ((abs(dl2bad[,22]) / abs(dl2bad[,21]))-1)
mean(dl2j2bad$Deadlift2Kg)

# When your third lift is BAD, how much percentage is it from the second?
dl3bad = df %>%
  filter(Name == Name1) %>%
  filter(Deadlift3Kg < 0)

dl3j2bad = 100 * ((abs(dl3bad[,23]) / abs(dl3bad[,22]))-1)
mean(dl3j2bad$Deadlift3Kg)
```

```{r}
###########################
# Attempts % when 3 for 3 #
###########################

dlgood = df %>%
  filter(Name == Name1) %>%
  filter(Deadlift1Kg > 0) %>%
  filter(Deadlift2Kg > 0) %>%
  filter(Deadlift3Kg > 0) 

dljump1good = 100 * ((abs(dlgood[,22]) / abs(dlgood[,21]))-1)
dljump2good = 100 * ((abs(dlgood[,23]) / abs(dlgood[,22]))-1)

mean(dljump1good$Deadlift2Kg)
mean(dljump2good$Deadlift3Kg)
mean(dljump1good$Deadlift2Kg) + mean(dljump2good$Deadlift3Kg)
```


# Progression
```{r}
#####################
# Lifts Progression #
#####################

dfname = df %>%
  filter(Name == Name1)

coeff = 0.5
ggplot(dftotal, aes(x=Date))+
  geom_line(aes(y = Best3SquatKg, color = "Squat"))+
  geom_line(aes(y = Best3DeadliftKg, color= "Deadlift"))+
  geom_line(aes(y = Best3BenchKg/coeff, color = "Bench Press")) +
  #geom_line(aes(y = BodyweightKg, color = "Bodyweight")) +
  scale_y_continuous(name = "Squat / Deadlift", sec.axis = sec_axis(~.*coeff, name="Bench Press"))
  
```

```{r}
###########################
# Coefficient progression #
###########################

dfname = df %>%
  filter(Name == Name1)

coeff2 = 1/5
ggplot(dftotal, aes(x=Date))+
  geom_line(aes(y = Wilks, color="Wilks"))+
  geom_line(aes(y = Glossbrenner, color="Glossbrenner"))+
  geom_line(aes(y = Goodlift/coeff2, color="Goodlift")) +
  geom_line(aes(y = Dots, color="Dots"))+
  scale_y_continuous(name = "Wilks, Glossbrenner, Dots", sec.axis = sec_axis(~.*coeff, name="Goodlift"))
```




