---
title: "Part 2"
author: "Pedro Lins"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages
library(gapminder)
library(tidyverse)
library(ggplot2)
```

# Load your data

```{r}
# Upload data as a .csv file
ds = "openpowerlifting.csv" 
df = read_csv(ds)
head(df)
```

# Exploratory Data Analysis
# Lifters Analysis
The goal here is to create a DSaaS (Data Science as a Service) that people can choose a lifter from the database and see his statistics. 

initially, lets evaluate the general formula for attempts index
if a lifter misses a lift, it shows as a negative value
if a lifter performed it, positive
so we want to count how many positives a lifter had and divide it by all attempts he did
this gives the probability of a lifter performing a good lift
this can be drilled down to each lift (squat, bench press, deadlift)
and by time (is it changing through time?)


```{r}
##################
# Attempts Index #
##################

#########
# Squat #
#########

Name1 = "Pedro Lins"

# First squat
sq1 = df %>%
  filter(Name == Name1) %>%
  count(Squat1Kg > 0)
colnames(sq1) = c("Squat1", "sq1")

  if(sq1[1,1] == TRUE){
    sq1idx = data.frame(c(1))
  } else {sq1idx = sq1[2,2] / (sq1[1,2] + sq1[2,2])
    }
colnames(sq1idx) = c("sq1idx")


# Second Squat
sq2 = df %>%
  filter(Name == Name1) %>%
  count(Squat2Kg > 0)
colnames(sq2) = c("Squat2", "sq2")

if(sq2[1,1] == TRUE){
    sq2idx = data.frame(c(1))
  } else {sq2idx = sq2[2,2] / (sq2[1,2] + sq2[2,2])
    }
colnames(sq2idx) = c("sq2idx")


# Thrid Squat
sq3 = df %>%
  filter(Name == Name1) %>%
  count(Squat3Kg > 0)
colnames(sq3) = c("Squat3", "sq3")

if(sq3[1,1] == TRUE){
    sq3idx = data.frame(c(1))
  } else {sq3idx = sq3[2,2] / (sq3[1,2] + sq3[2,2])
    }
colnames(sq3idx) = c("sq3idx")

# Fourth Squat will not get into the analysis since it is technically an outlier

sq = (sq1idx + sq2idx + sq3idx)/3
colnames(sq) = c("sq")
sq
```

```{r}
###############
# Bench Press #
###############

# First Bench Press
bp1 = df %>%
  filter(Name == Name1) %>%
  count(Bench1Kg > 0)
colnames(bp1) = c("Bench1", "bp1")

  if(bp1[1,1] == TRUE){
    bp1idx = data.frame(c(1))
  } else {bp1idx = bp1[2,2] / (bp1[1,2] + bp1[2,2])
    }
colnames(bp1idx) = c("bp1idx")
bp1idx

# Second Bench Press
bp2 = df %>%
  filter(Name == Name1) %>%
  count(Bench2Kg > 0)
colnames(bp2) = c("Bench2", "bp2")

  if(bp2[1,1] == TRUE){
    bp2idx = data.frame(c(1))
  } else {bp2idx = bp2[2,2] / (bp2[1,2] + bp2[2,2])
    }
colnames(bp2idx) = c("bp2idx")
bp2idx

# Thrid Bench Press
bp3 = df %>%
  filter(Name == Name1) %>%
  count(Bench3Kg > 0)
colnames(bp3) = c("Bench3", "bp3")

  if(bp3[1,1] == TRUE){
    bp3idx = data.frame(c(1))
  } else {bp3idx = bp3[2,2] / (bp3[1,2] + bp3[2,2])
    }

colnames(bp3idx) = c("bp3idx")


# Fourth Bench press will not get into the analysis since it is technically an outlier

bp = (bp1idx + bp2idx + bp3idx)/3
colnames(bp) = c("bp")
bp
```

```{r}
############
# Deadlift #
############

# First Deadlift
dl1 = df %>%
  filter(Name == Name1) %>%
  count(Deadlift1Kg > 0)
colnames(dl1) = c("Deadlift1", "dl1")

  if(dl1[1,1] == TRUE){
    dl1idx = data.frame(c(1))
  } else {dl1idx = dl1[2,2] / (dl1[1,2] + dl1[2,2])
    }
colnames(dl1idx) = c("dl1idx")
dl1idx

# Second Deadlift
dl2 = df %>%
  filter(Name == Name1) %>%
  count(Deadlift2Kg > 0)
colnames(dl2) = c("Deadlift2", "dl2")

  if(dl2[1,1] == TRUE){
    dl2idx = data.frame(c(1))
  } else {dl2idx = dl2[2,2] / (dl2[1,2] + dl2[2,2])
    }
colnames(dl2idx) = c("dl2idx")
dl2idx

# Thrid Deadlift
dl3 = df %>%
  filter(Name == Name1) %>%
  count(Deadlift3Kg > 0)
colnames(dl3) = c("Deadlift3", "dl3")

  if(dl3[1,1] == TRUE){
    dl3idx = data.frame(c(1))
  } else {dl3idx = dl3[2,2] / (dl3[1,2] + dl3[2,2])
    }
colnames(dl3idx) = c("dl3idx")
dl3idx

# Fourth Bench press will not get into the analysis since it is technically an outlier

dl = (dl1idx + dl2idx + dl3idx)/3
colnames(dl) = c("dl")
dl
```

```{r}
#################
# Overall Index #
#################

#index = (sq + bp + dl)/3
sq_index = data.frame(sq1idx, sq2idx, sq3idx, sq)
sq_trans <- as.data.frame(t(sq_index))
sq_trans
bp_index = data.frame(bp1idx, bp2idx, bp3idx, bp)
bp_trans<- as.data.frame(t(bp_index))
bp_trans
dl_index = data.frame(dl1idx, dl2idx, dl3idx, dl)
dl_trans<- as.data.frame(t(dl_index))
dl_trans

attempts = c(1,2,3,"overall")

newdf = data.frame(attempts, sq_trans, bp_trans, dl_trans)
colnames(newdf) = c("Attempts", "Squat", "Bench", "Deadlift")
print(newdf)

ggplot(newdf, aes(y = Squat, x = attempts))+
         geom_col()

```

